% Created 2023-04-17 Mon 18:42
% Intended LaTeX compiler: pdflatex
\RequirePackage{rotating}
\documentclass[12pt]{amsart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage{wasysym}
\newcommand{\Cov}{\ensuremath{\mbox{Cov}}}
\renewcommand{\Pr}{\ensuremath{\mbox{Pr}}}
\newcommand{\Eq}[1]{(\ref{eq:#1})}
\usepackage{bm}\usepackage{econometrics}
\usepackage{breqn}
\newcommand{\T}{\top}
\newtheorem{proposition}{Proposition} \newcommand{\Prop}[1]{Proposition \ref{prop:#1}}
%\newtheorem{problem}{Problem} \newcommand{\Prob}[1]{Problem \ref{prob:#1}}
%\newtheorem{theorem}{Theorem} \newcommand{\Thm}[1]{Theorem \ref{thm:#1}}
%\newtheorem{corollary}{Corollary} \newcommand{\Cor}[1]{Corollary \ref{cor:#1}}
%\newtheorem{remark}{Remark} \newcommand{\Rem}[1]{Remark \ref{rem:#1}}
%\newtheorem{condition}{Condition} \newcommand{\Cond}[1]{Condition \ref{cond:#1}}
%\newtheorem{lemma}{Lemma} \newcommand{\Lem}[1]{Lemma \ref{lem:#1}}
%\newtheorem{assumption}{Assumption} \newcommand{\Ass}[1]{Assumption \ref{ass:#1}}
\newcommand{\Fig}[1]{Figure \ref{fig:#1}} \newcommand{\Tab}[1]{Table \ref{tab:#1}}
\usepackage{dsfont}\newcommand{\one}{\ensuremath{\mathds{1}}}
\usepackage{xcolor}
\newcommand{\rv}[1]{\ensuremath{\textcolor{red}{#1}{}}}
%\newcommand{\rv}[1]{\ensuremath{{}_{rv}{#1}{}}}
%\newcommand{\rv}[1]{\ensuremath{\underline{#1}{}}}
\newcommand{\rvy}{\rv{y}}
\newcommand{\rvX}{\rv{X}}
\newcommand{\rvx}{\rv{x}}
\newcommand{\rvu}{\rv{u}}
\newcommand{\do}[1]{\ensuremath{\mbox{do}(#1)}}
\renewcommand{\E}{\ensuremath{\mathds{E}}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\author{Ethan Ligon}
\date{\today}
\title{Assignment 3 (Part A)}
\hypersetup{
 pdfauthor={Ethan Ligon},
 pdftitle={Assignment 3 (Part A)},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 29.0.50 (Org mode 9.6.1)}, 
 pdflang={English}}
\usepackage[style=authoryear]{biblatex}
\addbibresource{/home/ligon/bibtex/main.bib}
\begin{document}

\maketitle
This is the first part of Assignment 3, focusing on GMM.  I'll distribute a part B once it becomes a little more clear how much additional material we'll be able to cover before the end of the semester.

The entire assignment (both parts A \& B) will be due on May 1st.  As with the previous assignment, you are strongly encouraged to work as a  team, and to turn in a single assignment for grading.   The principal deliverable you turn in should be a link to a \texttt{github} repository, and you should organize your teams so as to provide constructive criticism to other teams.
\section{Exercises (GMM)}
\label{sec:orgb1fc826}
When we approach a new estimation problem from a GMM perspective
there's a simple set of steps we can follow.
\begin{itemize}
\item Describe the parameter space \(B\);
\item Describe a function \(g_j(b)\) such that \(\E g_j(\beta)=0\);
\item Describe an estimator for the covariance matrix \(\E g_j(\beta)g_j(\beta)^\T\).

\begin{enumerate}
\item Explain how the steps outlined above can be used to construct an optimally weighted GMM estimator.
\item Consider the following models.  For each, provide a causal
diagram; construct the optimally weighted GMM estimator of
the unknown parameters (various Greek letters); and give an
estimator for the covariance matrix of your estimates.  If
any additional assumptions are required for your estimator
to be identified please provide these.
\begin{enumerate}
\item \(\E\rvy = \mu\); \(\E(\rvy-\mu)^2 = \sigma^2\); \(\E(\rvy-\mu)^3 = 0\).
\item \(\rvy = \alpha + \rvX\beta + \rvu\); with \(\E(\rvX^\T\rvu)=\E\rvu=0\).
\item \(\rvy = \alpha + \rvX\beta + \rvu\); with \(\E(\rvX^\T\rvu)=\E\rvu = 0\),
and \(\E(\rvu^2)=\sigma^2\).
\item \(\rvy = \alpha + \rvX\beta + \rvu\); with \(\E(\rvX^\T\rvu)=\E\rvu=0\),
and \(\E(\rvu^2)=e^{X\sigma}\).
\item \(\rvy = \alpha + \rvX\beta + \rvu\); with
\(\E(\rv{Z}^\T\rvu)=\E\rvu=0\) and \(\E \rv{Z}^\T \rvX = \mQ\).
\item \(\rvy = f(\rvX\beta) + \rvu\); with \(f\) a known scalar function and with
\(\E(\rv{Z}^\T\rvu)=\E\rvu=0\) and \(\E \rv{Z}^\T \rvX
         f'(\rvX\beta) = \mQ(\beta)\).  (Bonus question: where does this last restriction come from, and
what role does it play?)
\item \(\rvy = f(\rvX,\beta) + \rvu\); with \(f\) a known function and with
\(\E(\rv{Z}^\T\rvu)=\E\rvu=0\) and \(\E \rv{Z}^\T\frac{\partial
         f}{\partial \beta^\T}(\rvX,\beta) = \mQ(\beta)\).
\item \(\rvy^\gamma = \alpha + \rvu\), with \(\rvy>0\) and \(\gamma\) a
scalar, and \(\E(\rv{Z}^\T\rvu)=\E\rvu=0\) and
\(\E\rv{Z}^\T\begin{bmatrix}\gamma
         \rvy^{\gamma-1}\\-1\end{bmatrix} =\mQ(\gamma)\).
\end{enumerate}

\item For each of the models above write a data-generating process in \texttt{python}.  Your function
\texttt{dgp} should take as arguments a sample size \texttt{N} and a vector of
``true'' parameters \texttt{b0}, and return a dataset \((y,X)\).
\item Select the most interesting of the data generating processes you
developed, and using the code in \texttt{gmm.py} or \texttt{GMM\_class.py} (see
\url{https://github.com/ligonteaching/ARE212\_Materials/}) use data
from your \texttt{dgp} to analyze the finite sample performance of the
corresponding GMM estimator you've constructed.  Of particular
interest is the distribution of your estimator using a sample
size \(N\) and how this distribution compares with the limiting
distribution as \(N\rightarrow\infty\).
\end{enumerate}
\end{itemize}

\section{Breusch-Pagan Extended}
\label{sec:orgb6be17c}
Consider a linear regression of the form
\begin{equation}
   y = \alpha + \beta x + u,
\end{equation}
with \((y,x)\) both scalar random variables, where it is assumed that
(a.i) \(\E(u\cdot x) = \E u = 0\) and (a.ii) \(\E(u^2|x)=\sigma^2\).
\begin{enumerate}
\item The condition a.i is essentially untestable; explain why.
\item \textcite{breusch-pagan79} argue that one can test a.ii via an
auxiliary regression \(\hat{u}^2 = c + d x + e\), where the \(\hat{u}\)
are the residuals from the first regression, and the test of a.ii
then becomes a test of \(H_0:d=0\).   Describe the logic of
the test of a.ii.
\item Use the two conditions a.i and a.ii to construct a GMM version of
the Breusch-Pagan test.
\item What can you say  about the performance or relative merits of the
Bruesch-Pagan test versus your GMM alternative?
\item Suppose that in fact that \(x\) is distributed uniformly over the
interval \([0,2\pi]\), and \(\E(u^2|x)=\sigma^2(x)=\sigma^2\sin(2x)\), thus
violating a.ii.  What can you say about the performance of the
Breusch-Pagan test in this circumstance?  Can you modify your GMM
test to provide a superior alternative?
\item In the above, we've considered a test of a specific functional
form for the variance of \(u\).  Suppose instead that we don't have
any prior information regarding the form of \(\E(u^2|x)=f(x)\).
Discuss how you might go about constructing an extended version
of the Breusch-Pagan test which tests for \(f(x)\) non-constant.
\item Show that you can use your ideas about estimating \(f(x)\) to
construct a more efficient estimator of \(\beta\) if \(f(x)\) isn't
constant.  Relate your estimator to the optimal generalized least
squares (GLS) estimator.
\end{enumerate}

\section{Tests of Normality}
\label{sec:org7798f32}
Suppose we have a sample of iid observations \(x_1,x_2,\dots,x_N\); we
want to test whether these are drawn from a normal distribution.
Note the fact that the integer central moments of the normal
distribution satisfy
\begin{align*}
 \mbox{E} x &= \mu\\
 \mbox{E}(x-\mu)^m &= 0\qquad\text{$m$ odd}\\
 \mbox{E}(x-\mu)^m &= \sigma^m(m-1)!!\qquad\text{$m$ even,}
\end{align*}
where \(n!!\) is the double factorial, i.e., \(n!!=n(n-2)(n-4)\dots\).

\begin{enumerate}
\item Using the analogy principle, construct an estimator for the first
\(k\) moments of the distribution of \(x\).  Use this to define a
\(k\)-vector of moment restrictions \(g_N(\mu,\sigma)\) satisfying
\(\mbox{E} g_N(\mu,\sigma) = 0\) under the null hypothesis of normality.
\item What is the covariance matrix of the sample moment restrictions
(again under the null)?
I.e., what can be said about \(\mbox{E} g_j(\mu,\sigma) g_j(\mu,\sigma)^\T - \mbox{E} g_j(\mu,\sigma) \mbox{E} g_j(\mu,\sigma)^\T\)?
\item Using your answers to the previous two questions, suggest a
GMM-based test of the hypothesis of normality, taking \(k>2\).
\item Implement the test you've devised using \texttt{python}.  You may want
to use \texttt{scipy.stats.distributions.chi2.cdf} and \texttt{scipy.optimize.minimize}.
\item What can be said about the optimal choice of \(k\)?
\item Compare the GMM estimates of \((\mu,\sigma)\) to the maximum
likelihood estimates of these parameters.  Do they differ?  Why?
\end{enumerate}
\section{Logit}
\label{sec:org37b277d}
This problem is meant to help draw connections between GMM estimators and maximum likelihood estimators, with a particular focus on the 'logit' model.  

The development of a maximum likelihood estimator typically begins with an assumption that some random variable has a (conditional) distribution which is known up a \(k\)-vector of parameters \(\beta\).
Consider the case in which we observe \(N\) independent realizations of a Bernoulli random variable \(\rv{Y}\), with \(\mbox{Pr}(\rv{Y}=1|X) = \sigma(\beta^\T X)\), and \(\mbox{Pr}(\rv{Y}=0|X) = 1-\sigma(\beta^\T X)\).

\begin{enumerate}
\item Show that under this model \(\E(\rv{Y}-\sigma(X\beta)|X)=0\).  Assume that \(\sigma\) is a known function, and use this fact to develop a GMM estimator of \(\beta\).  Is your estimator just- or over-identified?

\item Show that the likelihood conditional on realizations of data \((y,X)\) can be written as
\[    
            L(\beta|y,X) = \prod_{i=1}^N\sigma(\beta^\T X_i)^{y_i}\left(1-\sigma(\beta^\T X_i)\right)^{1-y_i}.
       \]
\item To obtain the maximum likelihood estimator (MLE) one can chose \(b\) to maximize \(\log L(b|y,X)\). When the likelihood is well-behaved, the MLE estimator satisfies the first order conditions (also called the ``scores'') from this maximization problem, in which case this is called a ``type I'' MLE.  Let \(\sigma(z)=\frac{1}{1+e^{-z}}\) (this is sometimes called the logistic function, or the sigmoid function), and obtain the scores \(S_N(b)\) for this estimation problem.  Show that \(\E S_N(\beta) = 0\).  Demonstrate that these moment conditions can serve as the basis for a GMM estimator of \(\beta\), and compare this estimator to the GMM estimator you developed above.  Which is more efficient, and why?
\end{enumerate}


\printbibliography
\end{document}